---
title: "Quarto for teaching using slides format"
author: "Mine Çetinkaya-Rundel"
format: 
  revealjs:
    code-link: false
    chalkboard: true
    multiplex: true
    incremental: true
    self-contained: false
editor: visual
execute: 
  echo: true
  code-fold: false
editor_options: 
  chunk_output_type: console
---

# Introduction

In this [YouTube](https://www.youtube.com/watch?v=EbAAmrB0luA), Dr. Mine Çetinkaya-Rundel demonstrates how she uses `Quarto` for teaching purposes.

## Data

For this analysis we will use penguins data set from the [palmerpinguins](https://CRAN.R-project.org/package=palmerpenguins) package.

```{r}
#| label: load-packages
#| message: false 
#| include: false

library(tidyverse)
library(ggthemes)
library(palmerpenguins)
library(gt)
```

```{r glimpse_penguins}
#| message: false
#| echo: true

glimpse(penguins)
```

## Species

@fig-bill-dims-species is a scatter plot of species of penguins

Note use `echo:fenced` to reveal all the codes

```{r depicting_penguin_species}
#| label: fig-bill-dims-species
#| warning: false
#| fig-width: 8
#| fig-asp: 0.618
#| fig-alt: |
#|   A scatterplot of penguins' bill depth and length, colored by 
#|   species of penguins. There is a relatively strong linear
#|   association.
#| fig-cap: |
#|   A scatterplot of penguins' bill depth and length, colored by 
#|   species of penguins.
#| code-line-numbers: "|7|8-9"
#| output-location: slide
#| echo: true #Use echo: fenced to reveal all coding

palmerpenguins::penguins |> 
  ggplot(aes(x = bill_length_mm,
           y = bill_depth_mm,
           color = species,
           shape = species)) +
  geom_point() +
  scale_color_colorblind() +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(x = "Bill length (mm)", 
       y = "Bill depth (mm)")

```

## Penguins

@tbl-penguins-top10 shows the fist 10 penguins from the dataset

```{r}
#| echo: true
#| label: tbl-penguins-top10
#| tbl-cap: First 10 penquins
penguins %>% # <1>
  slice_head(n = 10) %>% # <2>
  select(species, island, bill_length_mm,bill_depth_mm) %>% # <3>
  gt() # <3>
```
