---
title: "Quarto for teaching using slides format"
author: "Mine Çetinkaya-Rundel"
date: May 22, 2023
format:
  html:
    code-link: true
editor: visual
---

# Introduction

In this [YouTube](https://www.youtube.com/watch?v=EbAAmrB0luA), Dr. Mine Çetinkaya-Rundel demonstrates how she uses `Quarto` for teaching purposes.

For this analysis we will use penguins data set from the [palmerpinguins](https://CRAN.R-project.org/package=palmerpenguins) package.

Say you want to teach coding to your students, one thing to do is to encourage your students to read the documentation for functions. One thing we could do is to modify YAML. Adding a new field called `code-link: TRUE`. This allows the functions to be hyper linked.

```{r}
#| label: load-packages
#| message: false
#| include: false

library(tidyverse)
library(ggthemes)
library(palmerpenguins)
library(gt)
```

You can add options to executable code like this.

```{r glimpse_penguins}
#| echo: true
#| warning: false

glimpse(penguins)
```

## Species

@fig-bill-dims-species is a scatter plot of species of penguins

```{r}
#| warning: false
#| label: fig-bill-dims-species
#| fig-width: 5
#| fig-asp: 0.618
#| fig-alt: |
#|   A scatterplot of penguins' bill depth and length, colored by
#|   species of penguins. There is a relative strong, linear
#|   association.
#| fig-cap: |
#|   A scatterplot of penguins' bill depth and length, colored by
#|   species of penguins. 
#| code-line-numbers: "|23|24-25"
#| output-location: slide
#| echo: fenced #To frame the code

ggplot(data = penguins,
       aes(x = bill_length_mm,
           y = bill_depth_mm,
           color = species,
           shape = species)) +
  geom_point() +
  scale_color_colorblind() +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)")

```

## Penguins

@tbl-penguins-top10 shows the fist 10 penguins from the dataset

```{r}
#| echo: true
#| label: tbl-penguins-top10
#| tbl-cap: First 10 pinquins.

penguins %>% # <1>
  slice_head(n = 10) %>% # <2>
  select(species, island, bill_length_mm,bill_depth_mm) %>% # <3>
  gt() # <3>
```

1.  Start with the penguins data frame
2.  Then slice the top ten rows
3.  Select some variable and make a table with gt()
