---
title: "Quarto for publishing article"
author: "Mine Çetinkaya-Rundel"
format: 
  html:
    self-contained: true
    embed-resources: true
    toc: false
    code-fold: true
    theme: Zephyr
    number-sections: false
editor: visual
execute: 
  warning: false
  message: false
  echo: true
  output: true
editor_options: 
  chunk_output_type: console
---

# Introduction

In this [YouTube](https://www.youtube.com/watch?v=EbAAmrB0luA), Dr. Mine Çetinkaya-Rundel demonstrates how she uses `Quarto` for teaching purposes.

## Data

For this analysis we will use penguins data set from the [palmerpinguins](https://CRAN.R-project.org/package=palmerpenguins) package.

```{r}
#| label: load-packages
#| message: false 
#| include: false

library(tidyverse)
library(ggthemes)
library(palmerpenguins)
library(gt)
```

```{r glimpse_penguins}
#| message: false
#| echo: true

glimpse(penguins)
```

## Species

@fig-bill-dims-species is a scatter plot of species of penguins

```{r depicting_penguin_species}
#| label: fig-bill-dims-species
#| warning: false
#| fig-width: 8
#| fig-asp: 0.618
#| fig-alt: |
#|   A scatterplot of penguins' bill depth and length, colored by 
#|   species of penguins. There is a relatively strong linear
#|   association.
#| fig-cap: |
#|   A scatterplot of penguins' bill depth and length, colored by 
#|   species of penguins.
#| code-line-numbers: "7|8-9"
#| output-location: slide

palmerpenguins::penguins |> 
  ggplot(aes(x = bill_length_mm,
           y = bill_depth_mm,
           color = species,
           shape = species)) +
  geom_point() +
  scale_color_colorblind() +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(x = "Bill length (mm)", 
       y = "Bill depth (mm)")

```

## Penguins

@tbl-penguins-top10 shows the fist 10 penguins from the dataset. Notice we are adding annotations using `<>`

```{r}
#| echo: true
#| label: tbl-penguins-top10
#| tbl-cap: First 10 penquins

penguins |>  # <1>
  slice_head(n = 10) |>  # <2>
  select(species, island, bill_length_mm,bill_depth_mm) |>  # <3>
  gt() # <3>
```

1.  Start with the `penguins` data frame
2.  Then slide the top 10 rows
3.  Select some variables, and make a table with `gt()`
